% results.tex

\chapter{Results \& Discussion}

\section{Differentiation induces fibroblast-specific gene products in IMR90}

The transition from embryonic pluripotency to a specialized cell type naturally introduces a change in the expressed cell products.
Using Affymetrix microarrays, we observe that the overall trend towards differentiation marginally increases gene expression across
the entire proteome (mean = ?).  This observation is not surprising, and is thoroughly documented in the literature\cite{tuomela2012}
for various cell types.

We then sought to understand how expression changes were related to molecular function in the cell.  Using the ConsensusPathDB
tool\cite{kamburov2012}, we performed both over expression and gene ontology analysis of the 1\%  largest positively and negatively
and signalling pathways.  Furthermore, downregulated genes are involved in transcriptional regulation of pluripotent stem cells.

We asked whether transcriptional upregulation during differentiation could predict mutations, lesions, or break points seen in lung
cancer studies.  We acquired data for every type of lung cancer lesions from the Catalogue of Somatic Mutations in Cancer
(COSMIC)\cite{forbes2009} by gene.  Interestingly, the frequency of lesions reported by COSMIC was uncorrelated to those genes
expression changes (Pearsons $0.07$, $p < 1 \times 10^{-18}$.  There are a number of potential causes for the lack of signal in
our expression data.  Perhaps the expression arrays are not able to identify gene fusion products, a non-negligible portion of
the COSMIC mutation database.  More likely, as reported recently by Ashworth and colleagues, it is mutations in transcription factors
that cause gene expression changes in cancers, rather than in the genes themselves\cite{ashworth2014}.

\section{Contact maps of cell lines change drastically during differentiation}
% TODO

\section{ETC is conserved}


